\mytitle

### 2021-10-14 paper/noms-short-paper.tex
1.3 pages to go...

% !TeX document-id = {5530719d-34df-4dd8-b4b5-e6ed092c2b36}
% !TeX program = pdflatex
% !BIB program = biber
\documentclass[
%sigconf,natbib=false
%,anonymous=true
conference
,10pt
]{IEEEtran}

%\usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists]{savetrees}
%\usepackage[subtle, indent, lists]{savetrees}

@@

\usepackage{amsmath}
\let\openbox\relax
\usepackage{amsthm}
\newcommand\bmmax{2}
\usepackage{bm}

@@

@@

\usepackage[hang,flushmargin]{footmisc}

\begin{document}
	\title{Revisiting~the~Classics: Online~RL~in~the~Programmable~Dataplane}

@@

%	\author{Paper \# XXX, XXX pages}
	 \author{Kyle A. Simpson, Dimitrios P. Pezaros,\\University of Glasgow, Glasgow, Scotland}

%\setlength{\abovecaptionskip}{5pt}
%\setlength{\belowcaptionskip}{0pt}
%
%\setlength{\textfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
%\setlength{\floatsep}{3.0pt plus 0.0pt minus 2.0pt}
%
%\setlength{\dbltextfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
%\setlength{\dblfloatsep}{3.0pt plus 0.0pt minus 2.0pt}

@@

We present \emph{\approachshort{}}---\approach{}---the first work to bring \emph{online reinforcement learning} to the dataplane.
\approachshort{} makes online learning possible in constrained SmartNIC hardware by returning to classical RL techniques which simplify update logic and enable parallel processing.
Compared to offloading, we achieve a \SI{21}{\texttimes} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{\texttimes} improvement in online throughput for real-world policy designs.
In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, improving on offloading, and are quickly reconfigurable at runtime compared to reinstalling device firmware.

\end{abstract}

@@

%\vspace{-1em}
\section{Introduction}
\begin{figure}
	\centering
	\includegraphics[keepaspectratio, width=0.6875\linewidth]{figures/arch-with-p4}
	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane. SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible alongside P4 dataplanes. Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
\end{figure}
Automatic network optimisation, control, and defence are at last becoming commonplace.
Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing automatic traffic optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17,DBLP:conf/conext/GiladSGRS20}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
In RL methods, every change and its effects improve future decisions made by an agent.

@@

In parallel, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} (PDP) hardware~\parencite{DBLP:journals/micro/ZilbermanACM14, netronome-smartnic, xilinx-alveo, barefoot-intel} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
The promise of PDP hardware is that we can move the entire monitoring and analysis stack \emph{into the dataplane itself, and have it evolve to incorporate new approaches}.
Consider a DDoS mitigation system which could react to dataplane-specific, per-packet state in microseconds, and \emph{start learning again just as quickly} in the face of new data or behaviours.
The P4 ecosystem already presents novel, openly-available, fine-grained traffic measurement techniques that can be installed and controlled with ease~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
In addition, P4's control plane makes it easy to select which flows or packets are monitored in a live network and potentially allow control over traffic at the decision site.
As a result, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome,DBLP:journals/corr/abs-2002-08987} to take advantage of flow or per-packet state that cannot be efficiently processed or extracted at any other location in the network.
These works have shown the value of in-network ML: high-throughput, low latency response to network changes.
While they can exploit on-device state to provide reactive insight, the missing piece of the puzzle is learning and updating these ML analyses online without deferring to another machine in the network.
Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed by the community}.

It is important to make this feasible as offloading learning to commodity machines adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
For context, \emph{Deep Neural Network} (DNN) training relies on backpropagation to compute gradients, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many mini-batches of data for stable training.
These lead to high compute and storage costs, inducing capital expenditure for dedicated accelerators.
Moreover, high latencies \emph{caused} by offloading and expensive inference harm the learning process~\parencite{DBLP:journals/firai/TravnikMSP18} and impact runtime application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
If we can bring online learning \emph{to the dataplane}, then we can take advantage of rich, local state while minimising these latencies (and their impact on the learned policy).
This would also make it easy to train and prototype agent designs which can learn \emph{as the environment evolves}, or when there is too little data to model and simulate a problem.

@@

%The question we investigate is: can online RL be brought to the dataplane by returning to these computationally simpler methods, to act on locally extracted state?
%Can it be made more efficient \emph{by dataplane hardware}?
%Through this work, \approachshort{}, we can comfortably answer ``yes'' on both counts.
%Through \approachshort{}, w
We show that online RL can be brought to the dataplane by such methods, where it can act on locally extracted state and is made \emph{more efficient by dataplane hardware}.
In particular, we exploit how SmartNIC devices often expose general-purpose compute to provide path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
As many of these devices have engineering and development histories which predate P4, general compute beyond P4's limits~\parencite{p4-psa} is surprisingly common.
By executing on spare compute units, we prevent packet stalling and offer quick runtime reconfigurability.
%We look backwards to classical RL techniques (tile-coded policies, single-step RL algorithms), alongside the adaptations introduced by existing in-network ML (such as quantisation).
%Such adaptations unlock further improvements in parallel inference and update processing.
This paper contributes:
\begin{itemize}
	\item An analysis of why in-NIC RL is needed and best-placed to interact with the network, made feasible by classical RL methods and quantisation (\cref{sec:motivation}),
	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands of network traffic analysis (\cref{sec:design}),
	\item \emph{ParSa}, a wait-free, parallel, online RL algorithm to accelerate tile coded policy inference and updates (\cref{alg:parsa}),
	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{\texttimes} latency reduction compared to commodity hardware (\SI{21}{\texttimes} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
	\item A description of how \approachshort{} would integrate with state-of-the-art PDP applications to perform fully in-NIC, fast, automated DDoS mitigation (\cref{sec:potential-integrations}).
\end{itemize}

@@

\section{Preliminaries: towards in-NIC RL}\label{sec:motivation}
%\section{Preliminaries: RL for in-network computation}\label{sec:motivation}
By combining recent trends in PDP hardware with insights from the ML community (past and present), we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL and quantisation make this feasible.
Crucially, we focus on low port-density devices.
We envision that this processing is suited to the network edge or key vantage points, rather than core switches.
These devices' designs make it reasonable to move processing \emph{outside of the packet pipeline}; SmartNICs expose additional cores, and NetFPGAs allow for the synthesis of independent functional units.
%The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
The main benefit is that control logic can be moved onto the NIC \emph{without impacting packet processing}.
%?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
As Ethernet speed increases, packet deadlines grow tighter; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second} at \SI{40}{\giga\bit\per\second}, giving for example a \SI{3.99}{\micro\second} deadline on Netronome (which has \num{312} P4 pipelines).
In-pipeline execution thus increases the risk of packet drops or stalls.

@@

While P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to great interest in network programmability, it requires similar behaviour between device classes.
This is encoded by the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}---a lower bound on the device capabilities required to implement a P4 dataplane.
%We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
However, many compatible devices have legacies long predating these developments. 
For instance, many-core SoC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
%This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
Currently, low port-density devices like these are most likely to have specific general-purpose compute, as they are designed for high-performance offloading and middlebox development.
%Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
Regardless, floating-point operations key to ML workloads are very rarely supported outside of deep learning accelerators (which lack packet switching functionality).
Moreover, the limited memory and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.

@@

\emph{Reinforcement learning} (RL) algorithms train an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task~\parencite{RL2E}.
Like most ML methods, RL algorithms use gradient information to update the parameters used to approximate a function with \emph{reward measurements} from the environment.
RL's \emph{Markov Decision Process} (MDP) structure allows online learning of a state-action map in a model-free way, and can step through local minima when needed compared to ML.
In networking, this allows for learning to respond to on-device state or to handle rapidly evolving problems.
RL algorithms are generally agnostic to the policy approximation used, and classical single-step methods are simple---most of the compute work is occupied by the policy.
E.g., single-step, semi-gradient Sarsa~\parencite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to update and learn a policy online.

@@

\fakepara{RL in asynchronous environments}
%There remains some degree of divergence between the theory and implementation of RL agents.
%Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
%The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
%Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
Consider \cref{fig:state-slip}: the RL formulation assumes that an agent receives new state at fixed intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
This adds noise to the state-action mapping, which harms learning and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
%Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
Reducing these costs requires detailed agent-environment co-design.
This principle has influenced real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, where asynchrony is necessary for high-speed applications.
This often requires that state measurements are fused~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
`Stopping the world' causes significant performance loss, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control.

@@

This cost is what drives us to \emph{bring online reinforcement learning to the dataplane}---referring again to \cref{fig:state-slip}, we would place P4-based state measurement ($t_1$), simple policy approximation ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
In networks, actions and analysis are most likely to be installed on bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts---all of which make in-NIC training a perfect fit.
As such, our goal is to collocate \emph{all} functions which comprise an RL agent on the same chip or device.
Both programmable devices and the network environment make this more difficult, as we'll examine in the sequel.

@@

\subsection{Efficient policy approximation}
%Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
%This can then make \emph{online} learning of the task an attractive prospect, using a single available stream of experience.
%\emph{Deep neural networks} (DNNs), particularly when used as the basis for an RL agent's policy, require vast amounts of experience to converge on an accurate parameter set.
%In many problem domains, this equates to training from compute-years worth of distributed offline simulations, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
%Achieving stable learning requires sizeable batches of gradients to be computed, potentially using the entire experience replay from each simulation.
%Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the computational cost of gradient computation via backpropagation remains too significant on embedded hardware.
Complex \emph{Deep Neural Networks} (DNNs), particularly when used as an RL agent's policy, require vast amounts (possibly compute-years~\parencite{DBLP:journals/corr/abs-1912-06680}) of experience to converge on an accurate parameter set, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
Although there has been a vested interest in efficiently \emph{executing} complex function approximators like DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the storage and computational costs of backpropagation remain too significant on embedded hardware.

@@

We return to what classical RL methods can offer us; in particular, the simple linear function approximation given by \emph{tile coding}~\parencite[pp.\ \numrange{217}{221}]{RL2E}.
A policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
This allows any input-output mapping, such as converting packet header data into a preference list for queue priorities.
A state vector produces a single `hit' in each tiling, all of which then correspond to a list of action preference lists.
Inference (choosing an action) is simply summing over all such lists to obtain a final action preference list, selecting the maximum.
Updating uses this same list of lists after the next decision is made, adjusting the value of the last selected action using a \emph{temporal-difference value} ($\delta_t$) computed by Sarsa.
Crucially, this internal representation has no data dependencies between any tile calculations for an input.
As a result, both action selection and policy updates are map-reduce problems.
To select an action, we can produce a task for each tiling---retrieving the list of action values contributed by the tile activated by the input state.
An aggregate step sums up all action preference lists, and the action with the highest value is then selected.
Updating the policy (with a new $\delta_t$ value) has no aggregate step, and as tile indices map to disjoint memory regions no locks are required.
%As we show in \cref{sec:design}, it is possible to make the inference aggregate step wait-free.

>>>>

% !TeX document-id = {5530719d-34df-4dd8-b4b5-e6ed092c2b36}
% !TeX program = pdflatex
% !BIB program = biber
\documentclass[
%sigconf,natbib=false
%,anonymous=true
conference,
comsoc,
,10pt
]{IEEEtran}

\usepackage[subtle,bibnotes,charwidths,mathdisplays,indent,lists,tracking=normal]{savetrees}
%\usepackage[subtle, indent, lists]{savetrees}

@@

\usepackage{amsmath}
\let\openbox\relax
\usepackage{amsthm}
\newcommand\bmmax{2}
\usepackage[cmintegrals]{newtxmath}
\usepackage{bm}

@@

%\usepackage{xparse}
%\ExplSyntaxOn
%\NewDocumentCommand{\crefnames}{ m m m }
%{
%	\clist_map_inline:nn { #1 } { \crefname{##1}{#2}{#3} }
%}
%\ExplSyntaxOff
%
%\crefnames{section,subsection}{\S}{\S\S}
\crefformat{section}{\S{}#2#1#3}
\crefformat{subsection}{\S{}#2#1#3}

@@

\usepackage[hang,flushmargin]{footmisc}
\usepackage{orcidlink}

\newcommand{\mytitle}{Revisiting~the~Classics: Online~RL~in~the~Programmable~Dataplane}

\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black,
	pdftitle={\mytitle{}},
	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros}
}
\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}}}

\title{\mytitle}
\author{Kyle~A.~Simpson~\orcidlink{0000-0001-8068-9909}, Dimitrios~P.~Pezaros~\orcidlink{0000-0003-0939-378X}\\\emph{University of Glasgow}, Glasgow, Scotland\\\email{k.simpson.1@research.gla.ac.uk}, \email{Dimitrios.Pezaros@Glasgow.ac.uk}}

\begin{document}

@@

%	\author{Paper \# XXX, XXX pages}

\setlength{\abovecaptionskip}{0pt}
\setlength{\belowcaptionskip}{0pt}

\setlength{\textfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
\setlength{\floatsep}{3.0pt plus 0.0pt minus 2.0pt}

\setlength{\dbltextfloatsep}{3.0pt plus 0.0pt minus 3.0pt}
\setlength{\dblfloatsep}{3.0pt plus 0.0pt minus 2.0pt}

@@

We present \emph{\approachshort{}}---\approach{}---the first work to bring \emph{online reinforcement learning} to the dataplane.
\approachshort{} makes online learning possible in constrained SmartNIC hardware by returning to classical RL techniques---avoiding neural networks.
Our design allows weak yet highly parallel SmartNIC NPUs to be competitive against commodity x86 hosts, despite having fewer features and slower cores.
Compared to offloading, we achieve a \SI{21}{\texttimes} reduction in \num{99.99}\nthscript{th} tail inference times to \SI{34}{\micro\second}, and \SI{9.9}{\texttimes} improvement in online throughput for real-world policy designs.
In-NIC execution eliminates PCIe transfers, and our asynchronous compute model ensures minimal impact on traffic carried by a co-hosted P4 dataplane.
\approachshort{}'s design scales with additional resources at compile-time to improve upon both decision latency and throughput, and are quickly reconfigurable at runtime compared to reinstalling device firmware.
\end{abstract}

@@

%\vspace{-1em}
\section{Introduction}
\begin{figure}
	\centering
	\includegraphics[keepaspectratio, width=0.65\linewidth]{figures/arch-with-p4}
	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane. SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible alongside P4 dataplanes. Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
\end{figure}
Automatic network optimisation, control, and defence are at last becoming commonplace.
Adaptive techniques such as \emph{reinforcement learning} (RL) have led the charge in data-driven networking, enhancing automatic traffic optimisation~\parencite{DBLP:conf/sigcomm/ChenL0L18,DBLP:conf/sigcomm/MaoNA17}, congestion control~\parencite{DBLP:journals/corr/abs-1910-04054}, adaptive routing~\parencite{DBLP:conf/hotnets/ValadarskySST17,DBLP:conf/conext/GiladSGRS20}, resource management~\parencite{DBLP:conf/hotnets/MaoAMK16}, and packet classification~\parencite{DBLP:conf/sigcomm/LiangZJS19}.
In RL methods, every change and its effects improve future decisions.

@@

In parallel, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} (PDP) hardware~\parencite{DBLP:journals/micro/ZilbermanACM14, netronome-smartnic, xilinx-alveo, barefoot-intel} have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
The promise of PDP hardware is that we can move the entire monitoring and analysis stack \emph{into the dataplane itself, and have it evolve to incorporate new approaches}.
%Consider a DDoS mitigation system which could react to dataplane-specific, per-packet state in microseconds, and \emph{start learning again just as quickly} in the face of new data or behaviours.
The P4 ecosystem already presents novel, openly-available, fine-grained traffic measurement techniques that can be installed and controlled with ease~\parencite{DBLP:conf/sigcomm/GuptaHCFRW18,DBLP:conf/sigcomm/ChenFKRR18,DBLP:conf/sosr/GhasemiBR17}.
In addition, P4's control plane makes it easy to select which flows or packets are monitored in a live network and potentially allow control over traffic at the decision site.
As a result, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,DBLP:journals/corr/abs-2009-02353,langlet-ml-netronome,DBLP:journals/corr/abs-2002-08987} to take advantage of flow or per-packet state that cannot be efficiently processed or extracted at any other location in the network.
These works have shown the value of in-network ML: high-throughput, low latency response to network changes.
While they can exploit on-device state to provide reactive insight, the missing piece of the puzzle is learning and updating these ML analyses online without deferring to another machine in the network.
Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed by the community}.

It is important to make this feasible; offloading to commodity hosts adds PCIe delays~\parencite{DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/NeugebauerAZAL018}, but is required due to the complexity of modern ML.
For context, deep neural network training relies on backpropagation, can take vast amounts of offline simulation for RL~\parencite{DBLP:journals/corr/abs-1912-06680}, and needs many minibatches for stability.
These lead to high compute and storage costs, inducing capital expenditure for dedicated accelerators to overcome the batching needed to operate at line-rate---while BrainWave~\parencite{DBLP:conf/isca/FowersOPMLLAHAG18} can reduce batching (and thus tail latencies) by \qty{32}{\texttimes}~\parencite{Duarte2019}, inference still takes $\mathcal{O}{\left(\text{\si{\milli\second}}\right)}$.
Moreover, novel DMA techniques such as \emph{GPUDirect}~\parencite{gpudirect} halve but do not eliminate such PCIe transfers.
The high latencies \emph{caused} by offloading and inference harm learning~\parencite{DBLP:journals/firai/TravnikMSP18} and runtime application performance~\parencite{DBLP:journals/corr/abs-1910-04054}.
If we can bring online learning \emph{to the dataplane}, then we can take advantage of rich, local state while minimising these latencies (and their impact on the learned policy).
This would also make it easy to train and prototype agent designs which can learn \emph{as the environment evolves}, or when there is too little data to model and simulate a problem.

@@

%The question we investigate is: can online RL be brought to the dataplane by returning to these computationally simpler methods, to act on locally extracted state?
%Can it be made more efficient \emph{by dataplane hardware}?
%Through this work, \approachshort{}, we can comfortably answer ``yes'' on both counts.
%Through \approachshort{}, w
We show that online RL can be brought to the dataplane by such methods, where it can act on locally extracted state and is made \emph{more efficient by dataplane hardware}.
In particular, we exploit how SmartNIC devices often expose general-purpose compute~\parencite{DBLP:conf/sigcomm/LiuCSKPG19} to provide path-adjacent, on-chip RL in the dataplane (\cref{fig:netro-arch}).
As many of these devices have engineering and development histories which predate P4, general compute beyond P4's limits~\parencite{p4-psa} is surprisingly common.
By executing on spare compute units, we prevent packet stalling and offer quick runtime reconfigurability.
%We look backwards to classical RL techniques (tile-coded policies, single-step RL algorithms), alongside the adaptations introduced by existing in-network ML (such as quantisation).
%Such adaptations unlock further improvements in parallel inference and update processing.
This paper contributes:
\begin{itemize}
	\item An analysis of why in-NIC RL is needed and best-placed to interact with the network, made feasible by classical RL methods and quantisation (\cref{sec:motivation}),
	\item \emph{\approachshort{}}: a general-purpose in-NIC RL agent which scales with allocated device resources to meet latency or throughput demands of network traffic control (\cref{sec:design}),
	\item \emph{ParSa}, a wait-free, parallel RL algorithm to accelerate tile coded policy inference and updates (\cref{alg:parsa}),
	\item In-depth evaluation of how \approachshort{} affects carried dataplane traffic, performs under different policy sizes (simple/complex state), and improves on explicit offloading with a \SI{15}{\texttimes} latency reduction compared to commodity hardware (\SI{21}{\texttimes} for 99.99\nthscript{th} tail latencies) and an order of magnitude improvement in online throughput (\cref{sec:evaluation}).
	\item A description of how \approachshort{} would integrate with state-of-the-art PDP applications to perform fully in-NIC, fast, automated DDoS mitigation (\cref{sec:potential-integrations}).
\end{itemize}

@@

\section{Preliminaries: towards in-NIC RL}\label{sec:motivation}
%\section{Preliminaries: RL for in-network computation}\label{sec:motivation}
By combining recent trends in PDP hardware with insights from the ML community (past and present), we discuss why in-NIC RL is needed and best-placed to interact with the network, and how classical RL and quantisation make this feasible.
Crucially, we focus on low port-density devices.
%We envision that this processing is suited to the network edge or key vantage points, rather than core switches.
These devices' designs make it reasonable to move processing \emph{outside of the packet pipeline}; SmartNICs expose additional cores, and NetFPGAs allow for the synthesis of independent functional units.
%The main benefit of doing so is that core control logic can be moved as close to the device as possible \emph{without impacting packet processing rates}.
The main benefit is that RL logic can be moved onto the NIC \emph{without impacting packet processing}.
%?? Say more directly why this mode of operation is preferable to in-pipeline w/ P4
%As Ethernet speed increases, packet deadlines grow tighter; a \SI{64}{\byte} packet must be produced every \SI{12.8}{\nano\second} at \SI{40}{\giga\bit\per\second}, giving for example a \SI{3.99}{\micro\second} deadline on Netronome (which has \num{312} P4 pipelines).
%In-pipeline execution thus increases the risk of packet drops or stalls.

@@

While P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} has led to great interest in network programmability, it requires similar behaviour between device classes.
This is encoded by the \emph{Programmable Switch Architecture} (PSA)~\parencite{p4-psa}---a lower bound on the device capabilities required to implement a P4 dataplane.
%We would be remiss to believe that P4 and the PSA define all capabilities that programmable hardware supports.
However, many compatible devices long predate these developments. 
Many-core SoC-based Netronome~\parencite{netronome-smartnic}, NetFPGA SUME~\parencite{DBLP:journals/micro/ZilbermanACM14,DBLP:conf/fpga/IbanezBMZ19}, and other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo} allow virtually arbitrary programs to be specified and executed.
%This pattern extends to other SmartNICs~\parencite{nvidia-bluefield,xilinx-alveo}.
Currently, low port-density devices like these are most likely to have specific general-purpose compute and high degrees of parallelism~\parencite{DBLP:conf/sigcomm/LiuCSKPG19}, as they are designed for high-performance offloading and middlebox development.
%Even Intel Tofino ASICs~\parencite{barefoot-intel}, which architecturally mirror the PSA, expose additional matching capabilities and \emph{arithmetic logic unit} (ALU) functions via the Tofino Native Architecture.
%Regardless, floating-point operations key to ML workloads are very rarely supported outside of deep learning accelerators (which lack packet switching functionality).
%Moreover, the limited memory and per-packet timing constraints endemic to these devices make this class of in-NIC offloading challenging.

@@

\emph{Reinforcement learning} (RL) trains an \emph{agent} to choose an optimal sequence of actions in pursuit of a given task~\parencite{RL2E}.
Like most ML methods, RL algorithms use gradients to update the parameters used to approximate a function with \emph{reward measurements} from the environment.
RL's Markov Decision Process structure allows online learning of a state-action map in a model-free way, and can step through local minima when needed compared to ML.
In networking, this allows for learning from on-device state or handling rapidly evolving problems.
RL algorithms are agnostic to the policy approximation used so long as they are differentiable, and are computationally simple.
E.g., Sarsa~\parencite[pp. \numrange{217}{221}]{RL2E} requires only additions and multiplication to learn a policy online.

@@

%\fakepara{RL in asynchronous environments}
%%There remains some degree of divergence between the theory and implementation of RL agents.
%%Consider \cref{fig:state-slip}: the traditional formulation of a Markov decision process assumes that an agent receives a new view of the world's state at fixed time intervals, and then decides upon and executes an action instantly.
%%The reality is that state information takes time to traverse the network, service times are offset by how quickly hosts respond to interrupts and deserialise requests, and action preference lists are often computed via expensive policy approximations.
%%Action installation also incurs costs in fields such as network administration, initially to contact the controller and then for those actions to be installed via the control plane.
%Consider \cref{fig:state-slip}: the RL formulation assumes that an agent receives new state at fixed intervals, and then decides upon an action instantly, while in reality these incur transmission and compute delays.
%This adds noise to the state-action mapping, which harms learning and final accuracy even for simple grid world tasks~\parencite{DBLP:journals/firai/TravnikMSP18}.
%%Reordering algorithmic steps reduces these costs for online single-step algorithms, but reducing this further requires detailed agent-environment co-design.
%Reducing these costs requires detailed agent-environment co-design.
%This principle has influenced real network use cases, such RL-based congestion-control algorithms~\parencite{DBLP:journals/corr/abs-1910-04054}, where asynchrony is necessary for high-speed applications.
%This often requires that state measurements are fused~\parencite{DBLP:journals/corr/abs-1910-04054,DBLP:journals/tnsm/SimpsonRP20} while expensive computations are ongoing.
%`Stopping the world' causes significant performance loss, as inference takes up to \SI{30}{\milli\second} in the above work on congestion control.

@@

%This cost is what drives us to \emph{bring online reinforcement learning to the dataplane}---referring again to \cref{fig:state-slip}, we would place P4-based state measurement ($t_1$), simple policy approximation ($t_2$), and controlled systems ($t_3$) as close to one another as possible.
%In networks, actions and analysis are most likely to be installed on bump-in-the-wire NICs or middleboxes, and in the NICs of end-hosts---all of which make in-NIC training a perfect fit.
%As such, our goal is to collocate \emph{all} functions which comprise an RL agent on the same chip or device.
%Both programmable devices and the network environment make this more difficult, as we'll examine in the sequel.

@@

\subsection{Tile-coded policy approximation}
%Some problems either evolve over time in unpredictable ways, or cannot be easily modelled.
%This can then make \emph{online} learning of the task an attractive prospect, using a single available stream of experience.
%\emph{Deep neural networks} (DNNs), particularly when used as the basis for an RL agent's policy, require vast amounts of experience to converge on an accurate parameter set.
%In many problem domains, this equates to training from compute-years worth of distributed offline simulations, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
%Achieving stable learning requires sizeable batches of gradients to be computed, potentially using the entire experience replay from each simulation.
%Although there has been a vested interest in efficiently \emph{executing} more complex function approximators such as DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the computational cost of gradient computation via backpropagation remains too significant on embedded hardware.
%Complex \emph{Deep Neural Networks} (DNNs), particularly when used as an RL agent's policy, require vast amounts (possibly compute-years~\parencite{DBLP:journals/corr/abs-1912-06680}) of experience to converge on an accurate parameter set, which is at odds with the need to adapt to changes in the underlying problem \emph{as they happen}.
%Although there has been a vested interest in efficiently \emph{executing} complex function approximators like DNNs in-NIC~\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}, the storage and computational costs of backpropagation remain too significant on embedded hardware.

@@

Returning to classical RL, we examine the linear function approximation of \emph{tile coding}~\parencite[pp.\ \numrange{217}{221}]{RL2E}.
A policy is represented by sets of tiles, each covering one or more dimensions of the input state with several overlapping tilings (offset stepwise to provide generalisation).
%This allows any input-output mapping, such as converting packet header data into a preference list for queue priorities.
A state vector produces a single `hit' in each tiling, all of which then correspond to a list of action preference lists---where computing a hit requires division by a known tile width.
Inference is simply summing over all such lists to obtain a final preference list, selecting the highest-value action.
Learning uses this same list after the next decision is made, adjusting the value of the previous action using a \emph{temporal-difference value} ($\delta_t$) computed by Sarsa.
Crucially, this internal representation has no data dependencies between any tile calculations for an input.
%As a result, both action selection and policy updates are map-reduce problems.
%To select an action, we can produce a task for each tiling---retrieving the list of action values contributed by the tile activated by the input state.
%An aggregate step sums up all action preference lists, and the action with the highest value is then selected.
%Updating the policy (with a new $\delta_t$ value) has no aggregate step, and as tile indices map to disjoint memory regions no locks are required.
%As we show in \cref{sec:design}, it is possible to make the inference aggregate step wait-free.

%For instance, a policy with $m$ tiling sets (each having a set of input dimensions), with $n$ overlapping tilings then creates $m \times n$ separate tasks.
%As an example, a real-world policy size which we examine (\cref{sec:experimental-setup}) with one bias tile and \num{16} sets of \num{8} tilings creates \num{129} work items.
%Computing a tile hit uses one division per dimension, which is an expensive ALU operation to require.
%However, if tile widths are fixed at powers-of-two, then these may be replaced with right bitshifts.
%As we shall see shortly, this aligns with our choice of numerical representation.
